---
title: "Refinery Process Data Analysis in R | Yield & Properties"
output: html_document
---

# Refinery Process Data Analysis in R | Yield & Properties

## 1. Introduction

This reproducible R analysis explores process yield and properties. It is descriptive; observed associations are not causal evidence.

## 2. Analytical Questions

- What are the scale, structure and data-quality issues?
- Which patterns merit further investigation?

## 3. Data Loading

```{r}
find_csv <- function() { files <- c(if (dir.exists("/kaggle/input")) list.files("/kaggle/input", pattern="\\\\.csv$", recursive=TRUE, full.names=TRUE) else character(), list.files("../01_DADOS_BRUTOS", pattern="\\\\.csv$", full.names=TRUE)); if (!length(files)) stop("No CSV input found."); files[1] }
input_file <- find_csv()
data_raw <- read.csv(input_file, check.names=FALSE, stringsAsFactors=FALSE)
```

## 4. Data Overview

```{r}
dim(data_raw); utils::head(data_raw, 3); str(data_raw)
```

## 5. Data Quality

```{r}
empty_count <- function(x) if (is.character(x)) sum(trimws(x)=="", na.rm=TRUE) else 0L
quality_check <- data.frame(variable=names(data_raw), type=vapply(data_raw,function(x)class(x)[1],character(1)), missing_n=vapply(data_raw,function(x)sum(is.na(x)),integer(1)), empty_text_n=vapply(data_raw,empty_count,integer(1)))
quality_check; sum(duplicated(data_raw))
```

## 6. Data Preparation

```{r}
clean_name <- function(x) { x <- gsub("([a-z0-9])([A-Z])", "\\\\1_\\\\2", x, perl=TRUE); gsub("(^_|_$)","",gsub("[^a-z0-9]+","_",tolower(x))) }
data_clean <- data_raw; names(data_clean) <- clean_name(names(data_clean)); data_clean[] <- lapply(data_clean,function(x)if(is.character(x))trimws(x)else x)
```

## 7. Exploratory Data Analysis

```{r}
numeric_data <- data_clean[vapply(data_clean,is.numeric,logical(1))]
if(ncol(numeric_data)) print(summary(numeric_data))
if(ncol(numeric_data)>=2) print(round(cor(numeric_data,use="pairwise.complete.obs"),2))
```

## 8. Key Findings

All quantities are calculated directly from the input. Interpret rates and averages only after checking group sizes, units and data quality.

## 9. Technical Implications

The analysis supports prioritisation and investigation; it should not be used as causal evidence or an operational limit without domain review.

## 10. Limitations

Dataset provenance, definitions, sampling and licence status require verification before publication.

## 11. Conclusion

This is a portable, transparent starting point for Refinery Analytics exploration in R.

